\textbf{Predicting virus-receptor mutant binding by molecular dynamics simulation}

### 2013-06-20 paper.tex
Copy edits and final changes before sending back.

\newpage
\begin{abstract}
Existing computational methods to predict protein--protein interaction affinity often perform poorly in important test cases. In particular, the effects of multiple mutations, non-alanine substitutions, and flexible loops are difficult to predict with available tools and protocols. We present here a new method to interrogate affinity differences resulting from mutations in a host-virus protein--protein interface. Our method is based on extensive non-equilibrium all-atom simulations: We computationally pull the machupo virus (MACV) spike glycoprotein (GP1) away from the human transferrin receptor (hTfR1) and estimate affinity using the maximum applied force during a pulling simulation and the area under the force-versus-distance curve. We find that these quantities provide novel biophysical insight into the GP1/hTfR1 interaction. First, with no prior knowledge of the system we can differentiate among wild-type and mutant complexes. Second, although the static co-crystal structure shows two large hydrogen-bonding networks in the GP1/hTfR1 interface, our simulations indicate that only one of them is critical for the binding interaction. Third, one viral site known to be critical for infection may mark an important evolutionary suppressor site for infection-resistant hTfR1 mutants. Finally, our method provides an elegant framework to compare the effects of multiple mutations, individually and jointly, on protein--protein interactions.
\end{abstract}

\section*{Introduction}
The computational prediction of mutational effects on protein--protein interactions remains a challenging problem. Several methods are available to perform an energy difference calculation from an experimentally determined co-crystal structure. For example, computational alanine-scanning can be performed rapidly, with relatively low computational cost \citep{Grant2011,Kortemme2004}. However, such methods calculate the relative contribution to overall energy for each amino acid in a protein. They do not probe the interface directly, but rather calculate the total energy with each substitution \citep{Grant2011,Kortemme2004}. Alternative approaches have been developed using machine learning, training coefficients in a weighted equation containing geometric and energetic parameters \citep{Vreven2011,Vreven2012,Bajaj2011,Hwang2010}. Unfortunately, such machine-learning approaches often suffer in novel applications, for which available training sets are small or non-existent. As such, these methods are poorly suited for most host-virus protein--protein systems. By contrast, first principles methods offer the chance to forgo training. However, the currently available methods such as free energy perturbation (FEP) and thermodynamic integration (TI) rely on a two state model that phases directly from one three dimensional protein model to another with no intermediate transition states \citep{Gilson1997,Lu2004}. Thus, they can quickly lose physiological relevance as the two proteins become more dissimilar and as a result are largely limited to testing only sequential point mutants.

Here we propose a new method to probe protein--protein interactions directly, by pulling two proteins apart that are initially bound in a complex. We impart a pulling force within an all-atom molecular dynamics simulation on one member of the complex while the other is held in place. Then, we measure the force required for dissociation. As a proxy for relative binding affinity, we use the maximum applied force (MAF) required for separation and the area under the force-versus-distance curve (AUC).

@@

\subsection*{System Modeling}
We used the protein visualization software PyMOL \citep{PyMOL} to remove residues 121-190, 301-329, and 383-756 in the hTfR1. No residues were removed from the viral protein. We also removed 10 sugar residues from the native structure.

@@

\subsection*{Equilibration}
NAMD was used for all simulations in this study \citep{Phillips2005}. In addition to the modeled system, for equilibration we generated a configuration file that fixed the $\alpha$-carbon backbone. This was accomplished by setting $\beta = 1$. Further, we generated a configuration file with fixed $\alpha$-carbon atoms at residues 41-92 (numbered linearly, in this case, starting at 1 for the first amino acid as was required for VMD) in the hTfR1. The second file was used to affix a harmonic restraint, thus preventing any unfolding due to system reduction. More importantly, the harmonic restraint allowed the protein complex to equilibrate while preventing any drift from its predefined orientation. Finally, we calculated the system center and dimensions for use in molecular dynamics settings.

@@

\subsection*{Steered Molecular Dynamics}
Following equilibration, the final state of each simulation was used to generate a configuration file fixing the $\alpha$-carbon on residues 1, 58, 73-83, 96, 136, 137, 138, and 161 (again with linear numbering) in the hTfR1. Also, the $\alpha$-carbon of all residues (163-318 in linear numbering) in the GP1 were given an occupancy of 1, indicating they had an applied force during the simulation.

@@

\subsection*{Post-processing}
The python packages MDAnalysis \citep{Agrawal2011} and ProDy \citep{Bakan2011} were both used at various points in post-processing. The molecular trajectory (comprising the atomic coordinates per time) was parsed to compute the center-of-mass for each of the two complexes. The starting center-of-mass distance was set to zero and the distance was re-computed at each time step relative to the starting distance. The force is output directly to the standard log file.

@@

\subsection*{Steered molecular dynamics simulations}
To ensure a natural solution structure, we performed an initial equilibration phase to reduce the intrinsic crystallization distortions. First, we generated a physically realistic simulation world around the protein complex. The environment included an explicitly modeled bath of water, and a physiologically relevant concentration of ions. Next, with a fixed $\alpha$-carbon backbone we carried out successive rounds of energy minimization, followed by heating, and backbone release. Last,  with minimal restraints a short round of all-atom molecular dynamics \citep{Cuendet2008} was performed. We used the final state from each equilibrated system to restart another MD simulation. For SMD, at the centroid of several atoms a force was applied to a single point in GP1. By using a known spring constant, and pulling with a constant velocity, we could calculate the applied force during protein--protein dissociation \citep{Cuendet2008,Cuendet2011}. A typical averaged force curve comparison can be seen in Figure~\ref{fig:force_curve}. The force curves for each mutant were smoothed either by averaging over many replicates as in Figure~\ref{fig:force_curve} or by selecting an interpolation window for averaging. The latter approach was used for statistical analysis. In addition, to minimize the computational time in unproductive molecular dynamics, we shortened simulation time such that the interaction just barely dropped to zero near the end of each run. As seen in Figure~\ref{fig:force_curve}, this distance was relatively consistent among mutants. The supplementary movie visually illustrates the separation distance between peptide domains. The quantities MAF and AUC were derived from the force versus distances curves. Their summary statistics are reported in Table~\ref{tab:summary_statistics}. As we are more interested in the phenotypic impact of interface mutations we avoided many of the more physically rigorous, but technically complicated calculations that are possible with SMD \citep{Is2001A,Is2001B}.

@@

In total, we tested 7 point mutants and 3 double mutants in addition to the WT complex (Table~\ref{tab:summary_statistics}). The largest single-mutant impact was seen for a mutation to alanine at site Tyr211 in the hTfR1 ($ p < 0.001 $).  Several mutations at this site have proven capable of causing loss-of-infection, according to \textit{in vitro} flow-cytometry infection assays or known host-range limitations \citep{Rad2008,Rad20111,Rad20112}. Most likely, this effect is caused by the destruction of a critical hydrogen bond to Ser113 in the GP1. The lost hydrogen bond would lead to the subsequent loss of a large hydrogen-bonding network seen in the crystal structure \citep{Abraham2010}. In agreement with the static structure, our molecular dynamics simulations show Tyr211 is involved in a large hydrogen-bonding network (results not shown).

@@

We also simulated several point mutations at Asn348 in the hTfR1. Asn348Lys is reported in the literature to cause reduced infectivity \textit{in vivo} \citep{Rad2008,Abraham2010}. Further, in the crystal structure two residues on the GP1 and at least three on hTfR1 are involved in a fairly extensive hydrogen-bonding network. As discussed above, the alanine mutation at this site showed no significant difference in MAF or AUC from WT (Tables~\ref{tab:MAF_p} and ~\ref{tab:AUC_p}). In addition, the Asn348Lys mutation showed no significant difference from WT either, and neither did Asn348Trp. (For both of these mutations, mean MAF and mean AUC was lower than for WT, however. See Table~\ref{tab:summary_statistics})  On the other hand, there was a detectable difference between Asn348Ala and Asn348Lys (Tables~\ref{tab:MAF_p} and~\ref{tab:AUC_p}), with Asn348Lys being a weaker binder. Moreover, Asn348Trp showed nearly identical results to Asn348Lys. These results suggest that, in contrast to Tyr211, the critical difference at position 348 in hTfR1 is size exclusion rather than the destruction of a hydrogen-bonding network. The mutations to large amino acids (Asn348Trp and Asn348Lys) produced nearly identical affinity changes, whereas the mutations to amino acids not capable of hydrogen bonding (Asn348Ala and Asn348Trp) produced significantly different affinity changes (Table~\ref{tab:pairwise_differences}). To check the consistency of our results, we hypothesized that the combination of Tyr211Ala and Asn348Trp, being chemically disconnected in two different hydrogen-bonding networks, would lead to a synergistic loss-of-binding. As expected, the double mutant was the weakest binding mutant tested ($ p < 10^{-6} $) (Tables~\ref{tab:MAF_p} and~\ref{tab:AUC_p}) in this study. Further, according to MAF (but not AUC), the combination of Tyr211Ala and Asn348Trp also showed significantly weaker binding than Tyr211Ala by itself (Tables~\ref{tab:MAF_p} and~\ref{tab:AUC_p}). We suspect that the effect of Asn348Trp alone is near the limit of detection using our method. A larger number of replicates would possibly have resolved affinity differences between Asn348Trp and WT or other mutants more consistently.

Last, we further analyzed a single mutation in the GP1, vArg111Ala. It has been reported that vArg111Ala causes a loss-of-infection during \textit{in vitro} infectivity assays \citep{Rad20112}. As mentioned in the previous subsection, in our simulations this mutant showed no significant change in either MAF or AUC (Tables~\ref{tab:MAF_p} and~\ref{tab:AUC_p}), even though both quantities were, on average, lower than in WT (Table~\ref{tab:summary_statistics}). This result was somewhat surprising, since Tyr211Ala, presumably disrupting the same hydrogen-bonding network as vArg111Ala, displayed a significant reduction in affinity. To probe the interaction between position 111 in the GP1 and position 211 in the hTfR1 further, we also tested the double mutant vArg111Ala/Tyr211Ala. This double mutant showed affinity indistinguishable from WT and significantly higher than Tyr211Ala alone  (Table~\ref{tab:pairwise_differences}). Thus, vArg111Ala in the GP1 appeared capable of suppressing the loss-of-binding caused by Tyr211Ala. This result shows that the two sites do indeed interact significantly, and that replacing the hydrogen-bonding network at these sites with a hydrophobic interaction leads to comparable binding affinity.

@@

By extending the SMD technique to investigate a protein--protein interaction we have, in many ways, changed the scope of its application. Traditionally this technique is either applied to compute equilibrium quantities via a non-equilibrium approximation \citep{Park2003,Park2004,Buch2011,Giorgino2012}, used to estimate protein stability through unfolding \citep{Lu1999}, or used to calculate the absolute free energy of small molecule ligand binding \citep{Dixit2001}. Likewise, others have used SMD to understand the process of binding and unbinding at a resolution unmatched by experiment \citep{Cuendet2011,Giorgino2012}. And still others aimed at more esoteric problems apply driving forces to untie polypeptide knots in toy systems \citep{Sulkowska2010}. Here, we have shown that SMD can provide some insight into the \textit{relative} strength of protein--protein interactions. In contrast to more traditional absolute affinity calculations, relative binding is much more useful for experimental application. Further, our method complements whole cell infectivity assays, which demonstrate phenotypic effects but do not provide biochemical mechanism. Via SMD, one can separate mutations whose likely effect is altered binding affinity from mutations that may affect other factors such as protein expression, protein folding, or interactions with other proteins not considered in the simulation. Thus, SMD may open avenues for subsequent experimental work.

@@

Previous work has shown that the hTfR1 mutation Asn348Lys is capable of producing decreased infectivity \emph{in vitro} \citep{Rad2008,Rad20111}. Moreover, the molecular model based on x-ray diffraction data suggests destruction of a second hydrogen-bonding network as the cause \citep{Abraham2010}. In our simulations no individual point mutation at Asn348 showed significantly different affinity than the WT complex. Moreover, the mutation Asn348Ala resulted in an average MAF that was virtually identical to WT hTfR1 (Table~\ref{tab:pairwise_differences}). Thus, the putative hydrogen-bonding network involving Asn348 is probably a crystallization artifact. Even when paired with the highly destructive Tyr211Ala mutation, Asn348Ala was refractory to changes in binding affinity. By contrast, when Tyr211Ala was paired with large (Trp) and positively charged (Lys) substitutions at position 348, the result was a larger than expected synergistic difference. That is, the double mutant Tyr211Ala/Asn348Trp caused a much larger decrease in binding than we expected from either mutation individually. Moreover, although Asn348Trp and Asn348Lys individually failed to be distinguishable from WT hTfR1, they both caused a mean reduction in MAF and produced larger than expected synergistic effects. This data suggests that site 348 is size and charge restricted. Any mutation to a very large or positively charged side chain will likely produce decreased infectivity.

@@

Although our computational predictions largely agreed with the available literature, some mutations that are reported to cause reduced infectivity in experimental systems (most notably vArg111Ala and Asn348Lys, discussed above) were indistinguishable from the WT complex in our simulations. We would like to emphasize here that we cannot expect perfect agreement. While we have shown that our method can distinguish individual point mutations, we do not know the limit of detection with our method. It is possible that some mutants display measurable phenotypic effects in experiments yet appear identical under our method. More extensive sampling or refinement of the simulation protocol could help to differentiate such mutants (see also next paragraph). In addition, the available experimental data for the GP1/hTfR1 system were generally obtained from infection assays, not binding assays. There are several reasons other than binding due to which a viral mutation may cause a phenotypic difference in infectivity. First, from infectivity experiments it is not clear the extent to which a protein is grossly or partially misfolded. An additional analytical step with circular dichroism or an analogous technique may further clarify large-scale folding differences. Second, since we start with a bound structure, any changes that may dramatically affect the rate of association (different folds, trafficking issues, etc.) would be underestimated by our method. Third, although it is tempting to think diminished infectivity should correlate strongly with decreased binding affinity, it is likely that dramatically increased binding affinity would prohibit cellular membrane detachment after viral invasion. As a result, the virus could not get to its replication compartment (which is the cytosol in the case of Machupo) and would not reproduce.

@@

We have probed the biochemical effects of mutations at the host-virus interface between MACV and its human receptor. We applied an approach that imparts forces in an all-atom molecular dynamics (MD) simulation to drive the system through equilibrium. Our work has shown that with relatively modest computing resources it is possible to use MD simulations to investigate point mutations in the GP1/hTfR1 interface. In most cases, a double mutant was easily sufficient to resolve a difference in relative affinity. Moreover, relative to the WT complex one point mutant was capable of producing a large decrease in binding affinity. We applied our technique to the WT complex and 10 mutations in the GP1/hTfR1 interface. We were able to rationalize the results of infectivity data with a fundamental physical simulations. Our data has provided some direction for further biochemical intuition in this system. A polar residue is absolutely critical at position 211 in hTfR1. Any mutation to a nonpolar residue will probably lead to a rapid viral compensatory response. Also, although rodent populations may produce further mutations at Asn348 in hTfR1, unless a mutation is positively charged or very large it is unlikely to result in a large disruption to viral binding. We have also provided both a site (vArg111) and a mutation (vArg111Ala) that may be critical for future host-range adaptation of MACV, and thus should be monitored in surveillance efforts. Finally, our analysis offers a new means to systemically investigate host-virus interactions through computational molecular dynamics.

>>>>

\newpage

\begin{abstract}

Existing computational methods to predict protein--protein interaction affinity often perform poorly in important test cases. In particular, the effects of multiple mutations, non-alanine substitutions, and flexible loops are difficult to predict with available tools and protocols. We present here a new method to interrogate affinity differences resulting from mutations in a host-virus protein--protein interface. Our method is based on extensive non-equilibrium all-atom simulations: We computationally pull the machupo virus (MACV) spike glycoprotein (GP1) away from the human transferrin receptor (hTfR1) and estimate affinity using the maximum applied force during a pulling simulation and the area under the force-versus-distance curve. We find that these quantities provide novel biophysical insight into the GP1/hTfR1 interaction. First, with no prior knowledge of the system we can differentiate among wild-type and mutant complexes. Second, although the static co-crystal structure shows two large hydrogen-bonding networks in the GP1/hTfR1 interface, our simulations indicate that only one of them is critical for the binding interaction. Third, one viral site known to be critical for infection may mark an important evolutionary suppressor site for infection-resistant hTfR1 mutants. Finally, our method provides an elegant framework to compare the effects of multiple mutations, individually and jointly, on protein--protein interactions.
\end{abstract}

\section*{Introduction}

The computational prediction of mutational effects on protein--protein interactions remains a challenging problem. Several methods are available to perform an energy difference calculation from an experimentally determined co-crystal structure. For example, computational alanine-scanning can be performed rapidly, with relatively low computational cost \citep{Grant2011,Kortemme2004}. However, such methods calculate the relative contribution to overall energy for each amino acid in a protein. They do not probe the interface directly, but rather calculate the total energy with each substitution \citep{Grant2011,Kortemme2004}. Alternative approaches have been developed using machine learning, training coefficients in a weighted equation containing geometric and energetic parameters \citep{Vreven2011,Vreven2012,Bajaj2011,Hwang2010}. Unfortunately, such machine-learning approaches often suffer in novel applications, for which available training sets are small or non-existent. As such, these methods are poorly suited for most host-virus protein--protein systems. By contrast, first principles methods offer the chance to forgo training. However, the currently available methods such as free energy perturbation (FEP) and thermodynamic integration (TI) rely on a two state model that phases directly from one three dimensional protein model to another with no intermediate transition states \citep{Gilson1997,Lu2004}. Thus, they can quickly lose physiological relevance as the two proteins become more dissimilar and as a result are largely limited to testing only sequential point mutants.

Here we propose a new method to probe mutational effects in protein--protein interactions directly, by pulling two proteins apart that are initially bound in a complex. We impart a pulling force within an all-atom molecular dynamics simulation on one member of the complex while the other is held in place. Then, we measure the force required for dissociation. As a proxy for relative binding affinity, we use the maximum applied force (MAF) required for separation and the area under the force-versus-distance curve (AUC).

@@

\subsection*{System Modeling}

We used the protein visualization software PyMOL \citep{PyMOL} to remove residues 121-190, 301-329, and 383-756 in the hTfR1. No residues were removed from the viral protein. We also removed 10 sugar residues from the native structure.

@@

\subsection*{Equilibration}

NAMD was used for all simulations in this study \citep{Phillips2005}. In addition to the modeled system, for equilibration we generated a configuration file that fixed the $\alpha$-carbon backbone. This was accomplished by setting $\beta = 1$. Further, we generated a configuration file with fixed $\alpha$-carbon atoms at residues 41-92 (numbered linearly, in this case, starting at 1 for the first amino acid as was required for VMD) in the hTfR1. The second file was used to affix a harmonic restraint, thus preventing any unfolding due to system reduction. More importantly, the harmonic restraint allowed the protein complex to equilibrate while preventing any drift from its predefined orientation. Finally, we calculated the system center and dimensions for use in molecular dynamics settings.

@@

\subsection*{Steered Molecular Dynamics}

Following equilibration, the final state of each simulation was used to generate a configuration file fixing the $\alpha$-carbon on residues 1, 58, 73-83, 96, 136, 137, 138, and 161 (again with linear numbering) in the hTfR1. Also, the $\alpha$-carbon of all residues (163-318 in linear numbering) in the GP1 were given an occupancy of 1, indicating they had an applied force during the simulation.

@@

\subsection*{Post-processing}

The python packages MDAnalysis \citep{Agrawal2011} and ProDy \citep{Bakan2011} were both used at various points in post-processing. The molecular trajectory (comprising the atomic coordinates per time) was parsed to compute the center-of-mass for each of the two complexes. The starting center-of-mass distance was set to zero and the distance was re-computed at each time step relative to the starting distance. The force is output directly to the standard log file.

@@

\subsection*{Steered molecular dynamics simulations}
To ensure a natural solution structure, we performed an initial equilibration phase to reduce the intrinsic crystallization distortions. First, we generated a physically realistic simulation world around the protein complex. The environment included an explicitly modeled bath of water, and a physiologically relevant concentration of ions. Next, with a fixed $\alpha$-carbon backbone we carried out successive rounds of energy minimization, followed by heating, and backbone release. Last,  with minimal restraints a short round of all-atom molecular dynamics \citep{Cuendet2008} was performed. We used the final state from each equilibrated system to restart another MD simulation. For SMD, at the centroid of several atoms a force was applied to a single point in GP1. By using a known spring constant, and pulling with a constant velocity, we could calculate the applied force during protein--protein dissociation \citep{Cuendet2008,Cuendet2011}. A typical averaged force curve comparison can be seen in Figure~\ref{fig:force_curve}. The force curves for each mutant were smoothed either by averaging over many replicates as in Figure~\ref{fig:force_curve} or by selecting an interpolation window for averaging. The latter approach was used for statistical analysis. In addition, to minimize the computational time in unproductive molecular dynamics, we shortened the simulation time such that the interaction just barely dropped to zero near the end of each run. As seen in Figure~\ref{fig:force_curve}, this distance was relatively consistent among mutants. The supplementary movie visually illustrates the separation distance between peptide domains. The quantities MAF and AUC were derived from the force versus distances curves. Their summary statistics are reported in Table~\ref{tab:summary_statistics}. As we are more interested in the phenotypic impact of interface mutations we avoided many of the more physically rigorous, but technically complicated calculations that are possible with SMD \citep{Is2001A,Is2001B}.

@@

In total, we tested 7 point mutants and 3 double mutants in addition to the WT complex (Table~\ref{tab:summary_statistics}). The largest single-mutant impact was seen for a mutation to alanine at site Tyr211 in the hTfR1 ($ p < 0.001 $).  Several mutations at this site have proven capable of causing loss-of-infection, according to \textit{in vitro} flow-cytometry infection assays or known host-range limitations \citep{Rad2008,Rad20111,Rad20112}. Most likely, this effect is caused by the destruction of a critical hydrogen bond to Ser113 in GP1. The lost hydrogen bond would lead to the subsequent loss of a large hydrogen-bonding network seen in the crystal structure \citep{Abraham2010}. In agreement with the static structure, our molecular dynamics simulations show Tyr211 is involved in a large hydrogen-bonding network (results not shown).

@@

We also simulated several point mutations at Asn348 in the hTfR1. Asn348Lys is reported in the literature to cause reduced infectivity \textit{in vivo} \citep{Rad2008,Abraham2010}. Further, in the crystal structure two residues on GP1 and at least three on hTfR1 are involved in a fairly extensive hydrogen-bonding network. As discussed above, the alanine mutation at this site showed no significant difference in MAF or AUC from WT (Tables~\ref{tab:MAF_p} and ~\ref{tab:AUC_p}). In addition, the Asn348Lys mutation showed no significant difference from WT, and neither did Asn348Trp. (For both of these mutations, mean MAF and mean AUC was lower than for WT, however. See Table~\ref{tab:summary_statistics})  On the other hand, there was a detectable difference between Asn348Ala and Asn348Lys (Tables~\ref{tab:MAF_p} and~\ref{tab:AUC_p}), with Asn348Lys being a weaker binder. Moreover, Asn348Trp showed nearly identical results to Asn348Lys. These results suggest that, in contrast to Tyr211, the critical difference at position 348 in hTfR1 is size exclusion rather than the destruction of a hydrogen-bonding network. The mutations to large amino acids (Asn348Trp and Asn348Lys) produced nearly identical affinity changes, whereas the mutations to amino acids not capable of hydrogen bonding (Asn348Ala and Asn348Trp) produced significantly different affinity changes (Table~\ref{tab:pairwise_differences}). To check the consistency of our results, we hypothesized that the combination of Tyr211Ala and Asn348Trp, being chemically disconnected in two different hydrogen-bonding networks, would lead to a synergistic loss-of-binding. As expected, the double mutant was the weakest binding mutant tested ($ p < 10^{-6} $) (Tables~\ref{tab:MAF_p} and~\ref{tab:AUC_p}) in this study. Further, according to MAF (but not AUC), the combination of Tyr211Ala and Asn348Trp also showed significantly weaker binding than Tyr211Ala by itself (Tables~\ref{tab:MAF_p} and~\ref{tab:AUC_p}). We suspect that the effect of Asn348Trp alone is near the limit of detection using our method. A larger number of replicates would possibly have resolved affinity differences between Asn348Trp and WT or other mutants more consistently.

Last, we further analyzed a single mutation in GP1, vArg111Ala. It has been reported that vArg111Ala causes a loss-of-infection during \textit{in vitro} infectivity assays \citep{Rad20112}. As mentioned in the previous subsection, in our simulations this mutant showed no significant change in either MAF or AUC (Tables~\ref{tab:MAF_p} and~\ref{tab:AUC_p}), even though both quantities were, on average, lower than in WT (Table~\ref{tab:summary_statistics}). This result was somewhat surprising, since Tyr211Ala, presumably disrupting the same hydrogen-bonding network as vArg111Ala, displayed a significant reduction in affinity. To probe the interaction between position 111 in the GP1 and position 211 in the hTfR1 further, we also tested the double mutant vArg111Ala/Tyr211Ala. This double mutant showed affinity indistinguishable from WT and significantly higher than Tyr211Ala alone  (Table~\ref{tab:pairwise_differences}). Thus, vArg111Ala in the GP1 appeared capable of suppressing the loss-of-binding caused by Tyr211Ala. This result shows that the two sites do indeed interact significantly, and that replacing the hydrogen-bonding network at these sites with a hydrophobic interaction leads to comparable binding affinity.

@@

By extending the SMD technique to investigate a protein--protein interaction we have, in many ways, changed the scope of its application. Traditionally this technique is either applied to compute equilibrium quantities via a non-equilibrium approximation \citep{Park2003,Park2004,Buch2011,Giorgino2012}, used to estimate protein stability through unfolding \citep{Lu1999}, or used to calculate the absolute free energy of small molecule ligand binding \citep{Dixit2001}. Likewise, others have used SMD to understand the process of binding and unbinding at a resolution unmatched by experiment \citep{Cuendet2011,Giorgino2012}. And still others aimed at more esoteric problems apply driving forces to untie polypeptide knots in toy systems \citep{Sulkowska2010}. Here, we have shown that SMD can provide some insight into the \textit{relative} strength of protein--protein interactions. In contrast to more traditional absolute affinity calculations, relative binding is much more useful for experimental application. Further, our method complements whole cell infectivity assays, which demonstrate phenotypic effects but do not provide biochemical mechanisms. Via SMD, one can separate mutations whose likely effect is altered binding affinity from mutations that may affect other factors such as protein expression, protein folding, or interactions with other proteins not considered in the simulation. Thus, SMD may open avenues for subsequent experimental work.

@@

Previous work has shown that the hTfR1 mutation Asn348Lys is capable of producing decreased infectivity \emph{in vitro} \citep{Rad2008,Rad20111}. Moreover, the molecular model based on x-ray diffraction data suggests destruction of a second hydrogen-bonding network as the cause \citep{Abraham2010}. In our simulations no individual point mutation at Asn348 showed a significantly different affinity from the WT complex. Moreover, the mutation Asn348Ala resulted in an average MAF that was virtually identical to WT hTfR1 (Table~\ref{tab:pairwise_differences}). Thus, the putative hydrogen-bonding network involving Asn348 is probably a crystallization artifact. Even when paired with the highly destructive Tyr211Ala mutation, Asn348Ala was refractory to changes in binding affinity. By contrast, when Tyr211Ala was paired with large (Trp) and positively charged (Lys) substitutions at position 348, the result was a larger than expected synergistic difference. That is, the double mutant Tyr211Ala/Asn348Trp caused a much larger decrease in binding than we expected from either mutation individually. Moreover, although Asn348Trp and Asn348Lys individually failed to be distinguishable from WT hTfR1, they both caused a mean reduction in MAF and produced larger than expected synergistic effects. This data suggests that site 348 is size and charge restricted. Any mutation to a very large or positively charged side chain will likely produce decreased infectivity.

@@

Although our computational predictions largely agreed with the available literature, some mutations that are reported to cause reduced infectivity in experimental systems (most notably vArg111Ala and Asn348Lys, discussed above) were indistinguishable from the WT complex in our simulations. We would like to emphasize here that we cannot expect perfect agreement. While we have shown that our method can distinguish individual point mutations, we do not know the limit of detection with our method. It is possible that some mutants display measurable phenotypic effects in experiments yet appear identical under our method. More extensive sampling or refinement of the simulation protocol could help to differentiate such mutants (see also next paragraph). In addition, the available experimental data for the GP1/hTfR1 system were generally obtained from infection assays, not binding assays. There are several reasons other than binding due to which a viral mutation may cause a phenotypic difference in infectivity. First, from infectivity experiments it is not clear the extent to which a protein is grossly or partially misfolded. An additional analytical step with circular dichroism or an analogous technique may further clarify large-scale folding differences. Second, since we start with a bound structure, any changes that may dramatically affect the rate of association (different folds, trafficking issues, etc.) would be underestimated by our method. Third, although it is tempting to think diminished infectivity should correlate strongly with decreased binding affinity, it is likely that dramatically increased binding affinity would prohibit cellular membrane detachment after viral invasion. As a result, the virus would be prevented from reaching its replication compartment (which is the cytosol in the case of Machupo) and could not reproduce.

@@

We have probed the biochemical effects of mutations at the host-virus interface between MACV and its human receptor. We applied an approach that imparts forces in an all-atom molecular dynamics (MD) simulation to drive the system through equilibrium. Our work has shown that with relatively modest computing resources it is possible to use MD simulations to investigate point mutations in the GP1/hTfR1 interface. In most cases, a double mutant was easily sufficient to resolve a difference in relative affinity. Moreover, relative to the WT complex one point mutant was capable of producing a large decrease in binding affinity. We applied our technique to the WT complex and 10 mutations in the GP1/hTfR1 interface. We were able to rationalize the results of infectivity data with fundamental physical simulations. Our data has provided some direction for further biochemical intuition in this system. A polar residue is absolutely critical at position 211 in hTfR1. Any mutation to a nonpolar residue will probably lead to a rapid viral compensatory response. Also, although rodent populations may produce further mutations at Asn348 in hTfR1, unless a mutation is positively charged or very large it is unlikely to result in a large disruption to viral binding. We have also provided both a site (vArg111) and a mutation (vArg111Ala) that may be critical for future host-range adaptation of MACV, and thus should be monitored in surveillance efforts. Finally, our analysis offers a new means to systemically investigate host-virus interactions through computational molecular dynamics.
